\documentclass[12pt]{ucdavisthesis}

% PLEASE READ THE MANUAL - ucdavisthesis.pdf (in the package installation directory)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%               LATEX COMMANDS FOR DOCUMENT SETUP                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{bookmark}
\usepackage[us,nodayofweek,12hr]{datetime}
\usepackage{graphicx}

\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{graphicx}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{tabularx}
\RequirePackage{wasysym}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{color}
\RequirePackage{nicefrac}

\RequirePackage{lineno}
\RequirePackage{mathtools}
\RequirePackage[bibstyle=authoryear,citestyle=authoryear-comp,
                date=year,
                maxbibnames=9,maxnames=5,maxcitenames=2,
                backend=biber,uniquelist=false,uniquename=false,
                % style=apa,
                sorting=nyt,
                hyperref=true]{biblatex}

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}

\newcommand{\graham}[1]{\todo[size=\scriptsize, color=red!50]{#1}}
\newcommand{\vince}[1]{\todo[size=\scriptsize, color=blue!50]{#1}}

\newcommand{\cf}{\emph{cf.} }
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\cov}{Cov}
\DeclareMathOperator{\T}{{\mathrm{T}}}
\newcommand{\vect}[1]{\mathbf{#1}}
\newcommand{\nssh}{SSH_n}
\newcommand{\nnn}{\nonumber}
\newcommand{\chapquote}[2]{\begin{quotation} \textit{#1} \end{quotation} \begin{flushright} - #2\end{flushright} }

\DeclareMathOperator{\flt}{\mathrm{flat}}

\newcommand{\msxa}{\mars_X}
\newcommand{\fsxa}{\venus_X}

\bibliography{biblio}

\hyphenation{dis-ser-ta-tion blue-print man-u-script pre-par-ing} %add hyphenation rules for words TeX doesn't know


%\renewcommand{\rightmark}{\scriptsize A University of California Davis\ldots \hfill Rev.~\#1.0 \quad Compiled: \currenttime, \today}
% a fancier running header that can be used with draftcls options

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%        DOCUMENT SETUP AND INFORMATION FOR PRELIMINARY PAGES          %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title          {Exploring the Effects of Linkage for Ancestry and Temporal Polygenic Adaptation}
%Exact title of your thesis. Indicate italics where necessary by underlining or using italics. Please capitalize the first letter of each word that would normally be capitalized in a title.

\author         {Vincent Scott Buffalo}
%Your full name as it appears on University records. Do not use initials.

\authordegrees  { \;\;}
%Indicate your previous degrees conferred.

\officialmajor  {Population Biology}
%This is your official major as it appears on your University records.

\graduateprogram{Population Biology Graduate Group}
%This is your official graduate program name. Used for UMI abstract.

\degreeyear     {2019}
% Indicate the year in which your degree will be officially conferred.

\degreemonth    {September}
% Indicate the month in which your degree will be officially conferred. Used for UMI abstract.

\committee{Graham M. Coop}{Charles H. Langley}{Michael Turelli}{David J. Begun}{}
% These are your committee members. The command accepts up to five committee members so be sure to have five sets of braces, even if there are empties.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\copyrightyear{2019}
\nocopyright

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\dedication{\textsl{To my father, Ray Buffalo, who during my childhood worked
six days a week, sometimes commuting over 120 miles on weekends to work extra
hours, all to pay for my education. I will never forget my father visiting me
in San Jose after a long work day in San Francisco just to teach me fractions
in café bookstore after I struggled to understand the teacher's explanation. At
this time, my love of science grew, in no small part thanks to his dedication,
and his occasional gifts like Linus Pauling's \emph{General Chemistry} and Richard
Feynman's Feynman's \emph{Lost Lecture: The Motion of Planets Around the Sun} for
Easter, alongside a few dark chocolate bunnies. I owe you more than you know,
Dad.  }}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\abstract{

  Since Darwin's characterization, adaptation has often been portrayed as a
  gradual, plodding process acting over geological timescales to shape
  organisms' phenotypes to their environment. The masterful theoretical work of
  early geneticists like R.A. Fisher, Sewall Wright, and J.B.S Haldane
  developed population genetic models to understand the dynamics of selection,
  the impact of population structure, and the effects of genetic drift.
  However, this depiction of adaptation as a slow process has shifted, as
  numerous researchers have found numerous examples of rapid adaptation in the
  wild. I have been fascinated by rapid adaptation and models of recent
  ancestry since starting my PhD, and have developed theory and statistical
  methods exploring these ideas in my dissertation. Uniting my three chapters
  is the influence of linkage. Linkage shapes patterns of identity by descent
  (IBD) sharing between recently related individuals on the X chromosome which
  I explore in Chapter 1 by developing theoretical expressions for the number
  and length of IBD segments shared on the X chromosome between recent
  relatives. Due to the inheritance pattern of the X chromosome and the fact
  that recombination only occurs along the entire length of the X chromosome in
  females, IBD segment sharing on the X is rather different than sharing on the
  autosomes.

  In the second chapter of my dissertation, linkage determines the impact
  heritable fitness variation has in affecting the allele frequency
  trajectories of neutral sites in the genome. I investigate this in Chapter 2
  by developing theoretical expressions for the temporal autocovariance
  experienced by neutral sites, as they become associated with sites that
  determine a polygenic trait under selection. I propose that these temporal
  covariances can be estimated using genome-wide temporal data. In Chapter 3, I
  develop the statistical methods to accurately estimate temporal covariances,
  and quantify the impact of linked selection in three experimental evolution
  populations: two \emph{Drosophila simulans} evolve-and-resequence populations
  adapting to a lab environment, and one artificial selection experiment in
  mice. Overall, we find linked selection has a significant impact on allele
  frequency changes genome-wide.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgments{
  First and foremost, I am deeply indebted to my advisor Graham Coop for his
  mentorship and guidance during my PhD. I came into this PhD broadly
  interested in evolution, statistics, and genetics, with little idea of how to
  do rigorous  science, nor how to become a good biologist and population
  geneticist. Graham was endlessly patient with me as I explored various
  sub-disciplines and grew distracted by other projects; ultimately he
  (correctly) predicted that my interests would solidify around what this
  dissertation is about. There were times over the course of my PhD program
  that were extraordinarily difficult on me. As I encountered challenges,
  Graham was incredibly supportive and understanding. I would always look
  forward to our scientific discussions during our weekly meetings to rekindle
  my spirit. Looking back, I learned more about life and myself in this period
  than at any other point in my life, and I owe Graham an incredible debt for
  being so patient and supportive. I hope to pay his mentorship efforts forward
  as my career grows.
  
I would also like to thank members of the Coop Lab I have had the pleasure of
working with and learning from: Simon Aeschbacher, Chenling Antelope, Jeremy
Berg, Gideon Bradburd, Yaniv Brandvain, Erin Calfee, Nancy Chen, Doc Edge,
Katie Ferris, Emily Josephs, Ivan Juric, Kristin Lee, Matt Osmond, Alisa
Sedghifar, Anita To, and Sivan Yair. I owe a debt of gratitude to Jeff
Ross-Ibarra for first getting me interested in population genetics through  his
REHAB journal club, and allowing me to work in his lab before grad  school. I
am grateful for countless excellent conversations and good advice  from members
of my qualifying exam and dissertation committees, Chuck Langley, Dave Begun,
Jeff Ross-Ibarra, Michael Turelli, Rasmus Nielsen, and Bruce Rannala.

Throughout my PhD, numerous colleagues have collaborated on or provided
guidance with particular chapters. I would like to thank our collaborator Steve
M. Mount for his ideas and contributions in Chapter 1, as well as Amy Williams,
Noah Rosenberg, and two anonymous reviewers for their feedback during the
revision process at \emph{Genetics}. I am deeply indebted to Nick Barton for
providing extraordinarily helpful feedback for Chapter 2, as well as Enrique
Santiago, Aneil Agrawal, Bill Hill, John Kelly, Tyler Kent, Sally Otto,
Jonathan Pritchard, Kevin Thornton, and two anonymous reviewers for feedback on
previous versions of the manuscript. Finally, for providing assistance with the
datasets we reanalyzed and/or helpful feedback on Chapter 3, I would like to
thank Christian Schl{\"o}tterer, John Kelly, Kimberly Hughes, Frank Chan,
Campbell Rolian, Nick Barton, Alan Bergland, and Dmitri Petrov.  My research
has been supported by an NSF Graduate Research Fellowship (1148897); I would
like to thank the NSF for the opportunity to conduct this research.

I owe my family for their patience and encouragement during this process: Mom,
Dad, Anne, Lisa, Violet, Lauren, Dalilah, Grammy and Papa, and Grandma and
Grandpa Lamb. Lastly, I would like to thank my partner Sarah Friedman for
endless support and being a lovely dissertation-bae. I promise you I'll make up
for the lost backpacking trips this season.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Each chapter can be in its own file for easier editing and brought in with the \include command.
% Then use the \includeonly command to speed compilation when working on a particular chapter.
%%% \includeonly{ucdavisthesis_example_Chap1}

\begin{document}

%\newcommand{\bibfont}{\singlespacing}
% need this command to keep single spacing in the bibliography when using natbib

%\bibliographystyle{unsrtnat}
%many other bibliography styles are available (IEEEtran, mla, etc.). Use one appropriate for your field.

\makeintropages %Processes/produces the preliminary pages

\include{chapter_1}
\include{chapter_2}
\include{chapter_3}


\begingroup
\setstretch{0.8}
\setlength\bibitemsep{0pt}

% citations from chapter 1
\nocite{kram2015}
\nocite{dplyr}
\nocite{ggplot2}
\nocite{R}
\nocite{Rossum:1995:PRM:869369}
\nocite{C}
\nocite{colorspace}
\nocite{purrr}
\nocite{tidyr}


\printbibliography
\endgroup


% The UMI abstract uses square brackets!
% \UMIabstract[The UMI abstract is submitted online as plain text. It no longer has a word limit or specific formatting. Therefore, the UMIabstract command is now deprecated.]

\end{document} 
